A Hybrid Ensemble Learning Approach to Star-Galaxy Classification

### 2015-05-08 paper/main.tex
update paper

One of the advantages of a probabilistic classification is
that the threshold can be adjusted to produce
a more complete but less pure sample,
or a less complete but more pure one.
To compare the performance of probabilistic classification techniques
with that of morphological separation,
which has a fixed completeness ($c_g = 0.9964$, $c_s = 0.7145$)
at a certain purity ($p_g = 0.9597$, $p_s = 0.9666$),
we adjust the threshold of probabilistic classifiers
until the galaxy completeness $c_g$ matches
that of morphological separtion
to compute the galaxy purity $p_{g}$ at $c_g=0.9964$.
Similarly, the star purity $p_{s}$ at $c_{s}=0.7145$
is computed by adjusting the threshold
until the star completeness of each classifier is equal to
that of morphological separation.

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/purity_mag.pdf}
  \caption{Purity as a function of the $i$-band magnitude
           as estimated by the kernel density estimation (KDE) method
           for the differential counts.
           The top panel shows the histogram with a bin size of 0.1 mag
           and the KDE for objects in the test set.
           The second panel shows the fraction of stars estimated by KDE
           as a function of magnitude.
           The bottom two panels compare
           the galaxy and star purity values for BMC, TPC, and 
           morphological separation as functions of magnitude.
           Results for BMC, TPC, and morphological separation are in
           blue, green, and red, respectively.
           The $1 \sigma$ confidence bands are estaimted by
           bootstrap sampling.}
  \label{fig:purity_mag}
\end{figure}

@@

We present in Table~\ref{table:metrics_cut}
the same six metrics for each method,
and highlight the best method for each metric.
Overall, the results obtained for the reduced data set are remarkable.
With a smaller training set, our training based methods, TPC and SOMc,
suffer a significant decrease in performance.
The performance of morphological separation and HB
is essentially unchanged from Table~\ref{table:metrics_all}
as they do not depend on the training data.
Without sufficient training data,
the advantage of combining the predictions of different classifiers
is more obvious.
Even WA, the simplest of combination techniques, outperforms
all individual classification techniques in four metrics,
AUC, $p_s$ at $c_s=0.7145$, $p_g$ at $c_g=0.9600$, and $p_s$ at $c_s=0.2500$.
%Interestingly, although TPC does not perform better
%than morphological separation,
%BoM still chooses TPC as the best model within each SOM cell
%and shows exactly the same performance metrics as TPC.
%Stacking and BMC techniques still outperform
%all individual classification techniques,
Alhtough BoM always chooses TPC as the best model
when we have a high-quality training set,
it now chooses various methods in different bins
and outperforms all base classifiers.
While the performance of the stacking technique is only slightly worse
than that of BMC when we have a high-quality training set,
stacking now fails to outperform morphological separation.
BMC shows an impressive performance and 
outperforms all other classification techniques in all six metrics.
Overall, the improvements are small but still significant
since these metrics are averaged over the full test data.

@@

In Figure~\ref{fig:purity_mag_cut_integrated}, we show
the overall galaxy and star purity values as functions of magnitude
for the integrated counts.
Compared to Figure~\ref{fig:purity_mag_integrated},
the drop in the performance of TPC is clear.
However, even when some classifers have been trained on
a significantly reduced training set,
BMC maintains a galaxy purity of 0.970 and a star purity of 1.0
up to $i \sim 24.5$, and it sill outperforms morphological separation
at fainter magnitudes $i \ga 21$.

@@

We thank the referee for a careful reading of the manuscript
and comments that improved this work.
We thank Ignacio Sevilla for helpful and insightful conversations.
We acknowledge support from the National Science Grant No.
\textcolor{red}{XXXX}.

>>>>

One of the advantages of a probabilistic classification is
that the threshold can be adjusted to produce
a more complete but less pure sample,
or a less complete but more pure one.
To compare the performance of probabilistic classification techniques
with that of morphological separation,
which has a fixed completeness ($c_g = 0.9964$, $c_s = 0.7145$)
at a certain purity ($p_g = 0.9597$, $p_s = 0.9666$),
we adjust the threshold of probabilistic classifiers
until the galaxy completeness $c_g$ matches
that of morphological separation
to compute the galaxy purity $p_{g}$ at $c_g=0.9964$.
Similarly, the star purity $p_{s}$ at $c_{s}=0.7145$
is computed by adjusting the threshold
until the star completeness of each classifier is equal to
that of morphological separation.

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/purity_mag.pdf}
  \caption{Purity as a function of the $i$-band magnitude
           as estimated by the kernel density estimation (KDE) method
           for the differential counts.
           The top panel shows the histogram with a bin size of 0.1 mag
           and the KDE for objects in the test set.
           The second panel shows the fraction of stars estimated by KDE
           as a function of magnitude.
           The bottom two panels compare
           the galaxy and star purity values for BMC, TPC, and 
           morphological separation as functions of magnitude.
           Results for BMC, TPC, and morphological separation are in
           blue, green, and red, respectively.
           The $1 \sigma$ confidence bands are estimated by
           bootstrap sampling.}
  \label{fig:purity_mag}
\end{figure}

@@

We present in Table~\ref{table:metrics_cut}
the same six metrics for each method,
and highlight the best method for each metric.
Overall, the results obtained for the reduced data set are remarkable.
With a smaller training set, our training based methods, TPC and SOMc,
suffer a significant decrease in performance.
The performance of morphological separation and HB
is essentially unchanged from Table~\ref{table:metrics_all}
as they do not depend on the training data.
Without sufficient training data,
the advantage of combining the predictions of different classifiers
is more obvious.
Even WA, the simplest of combination techniques, outperforms
all individual classification techniques in four metrics,
AUC, $p_s$ at $c_s=0.7145$, $p_g$ at $c_g=0.9600$, and $p_s$ at $c_s=0.2500$.
%Interestingly, although TPC does not perform better
%than morphological separation,
%BoM still chooses TPC as the best model within each SOM cell
%and shows exactly the same performance metrics as TPC.
%Stacking and BMC techniques still outperform
%all individual classification techniques,
Although BoM always chooses TPC as the best model
when we have a high-quality training set,
it now chooses various methods in different bins
and outperforms all base classifiers.
While the performance of the stacking technique is only slightly worse
than that of BMC when we have a high-quality training set,
stacking now fails to outperform morphological separation.
BMC shows an impressive performance and 
outperforms all other classification techniques in all six metrics.
Overall, the improvements are small but still significant
since these metrics are averaged over the full test data.

@@

In Figure~\ref{fig:purity_mag_cut_integrated}, we show
the overall galaxy and star purity values as functions of magnitude
for the integrated counts.
Compared to Figure~\ref{fig:purity_mag_integrated},
the drop in the performance of TPC is clear.
However, even when some classifiers have been trained on
a significantly reduced training set,
BMC maintains a galaxy purity of 0.970 and a star purity of 1.0
up to $i \sim 24.5$, and it sill outperforms morphological separation
at fainter magnitudes $i \ga 21$.

@@

We thank Ignacio Sevilla for helpful and insightful conversations.
We acknowledge support from the 
National Science Foundation Grant No.\ AST-1313415.
RJB acknowledges support as an Associate
within the Center for Advanced Study at the University of Illinois.
